# Interpretation {#Interpretation}


## Overview

Chapter on interpretation.

```{r, echo=FALSE, results="asis", cache=TRUE}
# add steps to generate object up to here and cache
```



## Marker genes

Identify marker genes by testing genes for differential expression between clusters.

We use the `findMarkers` implementation in `scran` [@Lun2016-dn], using a binomial test, which tests for genes that differ in the proportion of expressing genes (expressed vs. not expressed) between clusters. This is a more stringent test than the default t-tests, and tends to select genes that are easier to interpret and validate experimentally.

[TO DO: demonstrate other types of tests]

```{r marker_genes, message=FALSE}
library(scran)

# set gene names for plotting
rownames(spe) <- rowData(spe)$gene_name

# test for marker genes
markers <- findMarkers(spe, test = "binom", direction = "up")

# returns a list with one DataFrame per cluster
markers
```

```{r marker_genes_heatmap, fig.width=5, message=FALSE}
library(pheatmap)

# plot log-fold changes for one cluster over all other clusters
interesting <- markers[[1]]
best_set <- interesting[interesting$Top <= 5, ]
logFCs <- getMarkerEffects(best_set)

pheatmap(logFCs, breaks = seq(-5, 5, length.out = 101))
```

```{r marker_genes_expression, fig.width=7, message=FALSE}
library(scater)

# plot log-transformed normalized expression of top genes for one cluster
top_genes <- head(rownames(interesting))

plotExpression(spe, x = "label", features = top_genes)
```


