# Load data {#load_data}

```{r setup, echo=FALSE, results="asis"}
library(rebook)
chapterPreamble(cache = TRUE)
```


## Overview

In this chapter, we show how to load and inspect the example dataset used in the rest of the analysis pipeline chapters.

The example dataset is a subset of a 10x Genomics Visium dataset from healthy human brain samples from the dorsolateral prefrontal cortex (DLPFC) region, from our publication @Maynard2020-ke.

For more details on the dataset, see the [Human DLPFC workflow](#human_DLPFC_dataset) chapter.



## Install packages

Install required packages (if not already installed).

```{r unref-install_packages, eval=FALSE}
install.packages("remotes")
remotes::install_github("drighelli/SpatialExperiment")

install.packages("BiocManager")
BiocManager::install("ExperimentHub")
remotes::install_github("lmweber/STdata")
```



## Load data

Load the human DLPFC example dataset from the [STdata](https://github.com/lmweber/STdata) package. Note that this is a subset containing only one sample (sample 151673) from the full dataset. See [Human DLPFC workflow](#human_DLPFC_dataset) for details.

```{r load_data, message=FALSE}
library(ExperimentHub)
library(STdata)

ehub <- ExperimentHub()

spe <- load_data("human_DLPFC")
```



## Inspect object

The example dataset is currently provided as a `SingleCellExperiment` object. We can inspect this object as follows.

[TO DO: update to use `SpatialExperiment`, which also includes the `spatialCoords` slot.]

We see that the object contains 33,538 rows (genes), 3,639 columns (spots or spatial barcodes), one `assay` (`counts` containing raw counts), `rowData` contains columns describing the genes, and `colData` contains columns describing the spots (including `x_coord` and `y_coord` containing the x and y coordinates from the tissue slide).

```{r inspect_object, message=FALSE}
spe
class(spe)
dim(spe)
names(assays(spe))
head(rowData(spe), 3)
head(colData(spe), 3)
```


