# Preamble {-}

```{r "intro"}
#| echo: false
intro <- tryCatch(
    {
        description <- packageDescription("PrinciplesSTA")
        pkg <- description$Package
        version <- description$Version
        authors <- eval(parse(text = description$Authors))
        license <- description$License
        glue::glue(
            "**Package:** {pkg}<br/>\n", 
            "**Compiled:** {as.character(Sys.Date())}<br/>\n", 
            "**Package version:** {version}<br/>\n", 
            "**R version:** <b style='color:red;'>{R.version.string}</b><br/>\n", 
            "**BioC version:** <b style='color:red;'>{BiocManager::version()}</b><br/>\n", 
            "**License:** {license}<br/>"
        )
    }, 
    error = function(e) {"Local preview"}
)
```

`r intro`


## Docker image {-}

A `Docker` image built from this repository is available here: 

👉 [ghcr.io/lmweber/principlessta](https://ghcr.io/lmweber/principlessta) 🐳

::: {.callout-tip icon='true'}
## Get started now 🎉

You can get access to all the packages used in this book in < 1 minute, 
using this command in a terminal: 

```{sh "docker", filename="bash"}
#| eval: false
docker run -it ghcr.io/lmweber/principlessta:devel R
```

:::


## RStudio Server {-}

An RStudio Server instance can be initiated from the `Docker` image as follows: 

```{sh "rstudio", filename="bash"}
#| eval: false
docker run \
    --volume <local_folder>:<destination_folder> \
    -e PASSWORD=OHCA \
    -p 8787:8787 \
    ghcr.io/lmweber/principlessta:devel
```

The initiated RStudio Server instance will be available at 
[https://localhost:8787](https://localhost:8787).

<!-- 
# System dependencies and conda 

A pre-configured `micromamba` environment called `BiocBook` is available 
once the Docker image is loaded. It provides all the softwares listed in 
`requirements.yml`. 

```{sh "micromamba", filename="bash"}
#| eval: false
docker run -it ghcr.io/lmweber/principlessta:devel bash
micromamba activate BiocBook
``` 
-->


## Session info {-}

::: {.callout-note collapse="true"}

## Click to expand

```{r "session info"}
#| cache: false
sessioninfo::session_info(
    installed.packages()[,"Package"], 
    include_base = TRUE
)
```

:::


## BiocBook

The book is built using the [BiocBook](https://github.com/js2264/BiocBook) package, developed by [Jacques Serizay](https://jserizay.com/) for building Quarto books with Bioconductor. For more details on building a book with BiocBook, see the [BiocBook book](https://jserizay.com/BiocBookDemo/devel/).
