# Load data {#sec-load-data}


## Overview

In the following chapters, we use spatial transcriptomics datasets that have been formatted into `SpatialExperiment` or other Bioconductor data classes (see @sec-bioconductor-data-classes).

Here, we load a 10x Genomics Visium dataset that will be used in several of the following chapters.

This dataset has previously been preprocessed using data preprocessing procedures with tools outside R and formatted into `SpatialExperiment` format. For more details on data preprocessing procedures for the 10x Genomics Visium platform, see the related online book [Visium Data Preprocessing](https://lmweber.org/Visium-data-preprocessing/).

The dataset here is available in `SpatialExperiment` format from the [STexampleData](https://bioconductor.org/packages/STexampleData) Bioconductor package.



## Dataset

This dataset consists of one sample (Visium capture area) from one donor, consisting of postmortem human brain tissue from the dorsolateral prefrontal cortex (DLPFC) brain region, which was measured with the 10x Genomics Visium platform. The dataset is described in the original publication by @Maynard2021.

More details on the dataset are also included in @sec-workflow-human-dlpfc.

<!-- To do: move additional details on dataset here (from DLPFC workflow chapter) -->



## Load data

Load the dataset in `SpatialExperiment` format from the [STexampleData](https://bioconductor.org/packages/STexampleData) Bioconductor package.

```{r, message=FALSE}
library(SpatialExperiment)
library(STexampleData)

# load object
spe <- Visium_humanDLPFC()
```



## SpatialExperiment object

Check the contents of the `SpatialExperiment` object. For more details on the `SpatialExperiment` structure, see @sec-bioconductor-data-classes.

```{r, message=FALSE}
# check object
spe

# number of genes (rows) and spots (columns)
dim(spe)

# names of 'assays'
assayNames(spe)

# row (gene) data
head(rowData(spe))

# column (spot) data
head(colData(spe))

# spatial coordinates
head(spatialCoords(spe))

# image data
imgData(spe)
```
