# Load data {#sec-load-data}


## Overview

In the following analysis chapters, we use a pre-prepared dataset where we have previously applied data preprocessing procedures (using tools outside R and Bioconductor) and saved the object in the `SpatialExperiment` format. This is available from the [STexampleData](https://bioconductor.org/packages/STexampleData) package.

The dataset consists of a single sample of human brain from the dorsolateral prefrontal cortex (DLPFC) region, measured using the 10x Genomics Visium platform, sourced from @Maynard2021. The dataset is also described in more detail in @sec-workflow-human-dlpfc.



## Load data

Here, we show how to load the data from the `STexampleData` package.

```{r, message=FALSE}
library(SpatialExperiment)
library(STexampleData)

# load object
spe <- Visium_humanDLPFC()
```



## SpatialExperiment object

Next, we inspect the `SpatialExperiment` object. For more details, see @sec-bioconductor-data-classes.

```{r, message=FALSE}
# check object
spe

# number of features (rows) and spots (columns)
dim(spe)

# names of 'assay' tables
assayNames(spe)

# features metadata
head(rowData(spe))

# spot-level metadata
head(colData(spe))

# spatial coordinates
head(spatialCoords(spe))

# image metadata
imgData(spe)
```
