# Chapter title

## Preamble

### Introduction

> Start with some words on what motivates this chapter.

### Dependencies

> Should consist of 1-2 chunks that set up the computational environment (libraries, parallelizaton, seed, etc.) and (if applicable) load data from earlier chapter(s).

```{r loading}
set.seed(77)
library(ggplot2)
library(BiocParallel)
bp <- MulticoreParam(10)
se <- readRDS("obj.rds")
```

## 1st section

> Please stick to an 80-character limit in code chunks, and use `BiocStyle` when mentioning any Bioc/CRAN/GH packages.

## 2nd section

> To keep things concise, use a collapsed box when you see fit (e.g., more details, advanced content, etc.):

::: {.callout-note collapse="true" title="itbeme"}

Anything can go in here (text/math, code chunks)...

```r
# do some fancy things
```

:::

## Appendix

### Saving {.unnumbered}

> If applicable, save objects for downstream.

```{r saving}
saveRDS(obj, "xxx-foo.rds")
```

### Benchmarks {.unnumbered}

> Bullet list of benchmark papers, ideally giving a 1-2 sentence summary of what's being compared (e.g., how many methods/datasets, which technologies, etc.).

### References {.unnumbered}

> Auto-populated; add nothing.
